Demo arXiv template

### 2022-10-27 paper/paper.qmd
restructure section 2 data pipeline in R

---
title: "Demo arXiv template"
format:
  arxiv-pdf:
    keep-tex: true  
  arxiv-html: default
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
    url: https://huizezhangsh.netlify.app/
  - name: Collaborators
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
abstract: |
  - Indices, useful, quantify severity, early monitoring, 
  - A huge number of indices have been proposed by domain experts, however, a large majority of them are not being adopted, reused, and compared in research or in practice. 
  - One of the reasons for this is the plenty of indices are quite complex and there is no obvious easy-to-use implementation to apply them to user's data. 
  - The paper describes a general pipeline framework to construct indices from spatio-temporal data, 
  - This allows all the indices to be constructed through a uniform data pipeline and different indices to vary on the details of each step in the data pipeline and their orders. 
  - The pipeline proposed aim to smooth the workflow of index construction through breaking down the complicated steps proposed by various indices into small building blocks shared by most of the indices.
  - The framework will be demonstrated with drought indices as examples, but appliable in general to environmental indices constructed from multivariate spatio-temporal data
keywords: 
  - spatio-temporal data
  - indices
  - data pipeline
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: inline
---

@@

The concept of pipeline (and pipe) is not a new concept to computing. It refers to a set of data processing elements connected in series, where the output of one element is the input of the next one.
@wickham_plumbing_2009 argues that whether made explicit or not, pipeline has to be presented in every graphics program.
The paper also argues that breaking down graphic rendering into steps is beneficial for understand the implementation and compare between different graphic systems.

*The pipeline system would work partially because of an agreed data structure (tidy data) and [same input and output format] - why the pipeline approach would work.*

The tidyverse suite provides a collection of R package for conducting data analysis.
These tools directly operate on data, and users, even new users, can easily read from the code on what operation is performed on the data and what to expect as the output.
All of these are built upon a simple prior knowledge of the tidy data principle [@wickham_tidy_2014], which prescribes three rules to structure tabular data: 1) Each variable forms a column, 2) Each observation forms a row, and; 3) Each type of observational unit forms a table.

Constructing indices also involves this working pattern of trials and errors on different options before finding the final solution. A tidy workflow will give researchers more flexibility to tune the steps as they wish and while maintain a readable and intuitive code base.

To better understand the pipeline approach in statistic computing, examples in interactive graphics and machine learning models are given.
In interactive graphics, the data pipeline is used to describe how the data is processed to be rendered on the screen. 
Examples of early work on building pipeline for interactive graphics includes @buja_elements_1988, @sutherland_orca_2000, and @xie_reactive_2014. 
Steps in these pipelines include non-linear transformation, variable standardization, randomization and dimension reduction.

A more familiar example from tidymodels [@tidymodels] presents a set of packages for machine learning models following the tidyverse principles [@tidyverse]. 
The machine learning pipeline constitutes data resampling, feature engineering, model fitting, model tuning, and model evaluation each being responsible by a separate package. 
Before tidymodel, analysts who wish to fit different models need to learn the workflow of each individual algorithm and need to rewrite most of the model fitting, tuning, and evaluation code when switching from one model to another. 
The tidymodel approach unifies a wide range of models under a single pipeline where users can substitute different models while maintain the same pipeline.
This allows users to easily experiment with different machine learning models a lot easier.

Both arguments are applicable to index construction.
Papers that proposing new indices always present a workflow diagram to illustrate how data is transformed into the final indices, but not much implementation has explicitly present this this step-by-step transformation in the code design.

>>>>

---
title: "Demo arXiv template"
format:
  arxiv-pdf:
    keep-tex: true  
  arxiv-html: default
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
    url: https://huizezhangsh.netlify.app/
  - name: Collaborators
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
abstract: |
  - Indices, useful, quantify severity, early monitoring, 
  - A huge number of indices have been proposed by domain experts, however, a large majority of them are not being adopted, reused, and compared in research or in practice. 
  - One of the reasons for this is the plenty of indices are quite complex and there is no obvious easy-to-use implementation to apply them to user's data. 
  - The paper describes a general pipeline framework to construct indices from spatio-temporal data, 
  - This allows all the indices to be constructed through a uniform data pipeline and different indices to vary on the details of each step in the data pipeline and their orders. 
  - The pipeline proposed aim to smooth the workflow of index construction through breaking down the complicated steps proposed by various indices into small building blocks shared by most of the indices.
  - The framework will be demonstrated with drought indices as examples, but appliable in general to environmental indices constructed from multivariate spatio-temporal data
keywords:
  - indices
  - data pipeline
  - software design
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: inline
---

@@

**Why you should care about pipeline**

The concept of pipeline (and pipe) is not a new concept to computing. It refers to a set of data processing elements connected in series, where the output of one element is the input of the next one.
@wickham_plumbing_2009 argues that whether made explicit or not, pipeline has to be presented in every graphics program.
The paper also argues that breaking down graphic rendering into steps is beneficial for understand the implementation and compare between different graphic systems. 
The discussion on pipeline construction is well documented in early interactive graphics software: @buja_elements_1988, @sutherland_orca_2000, and @xie_reactive_2014 and their pipeline steps include non-linear transformation, variable standardization, randomization and dimension reduction.

**What is pipeline, its underlying software design philosophy, and how these are reflected in R**

One of the most commonly known pipeline example is perhaps the Unix pipeline where programs can be concatenated with `|` to flow the output from the last program into the next program. 
The pipeline design avoid writing a complicated program to solve a complex problem but building simple programs that do one thing well and work well together.
This design philosophy is also embraced in the R community by the tidyverse principles. 
In tidyverse, the flow from the previous program to the next is ensured through reusing the tidy data format [@wickham_tidy_2014], which prescribes three rules on how to layout tabular data.
To solve a complicated data problem using tidyverse, one typically builds the solution using a collection of functions and these functions have a strong human centered design with intuitive naming and easy to reference through autocomplete.

**A proper pipeline example in R: tidymodel**

A pipeline example in R that adopts the tidyverse principle is the tidymodel suite [@tidymodels] for building machine learning models. 
The machine learning pipeline constitutes data resampling, feature engineering, model fitting, model tuning, and model evaluation each being responsible by a separate package. 
Before tidymodel, analysts who wish to fit different models need to learn the workflow of each individual algorithm and need to rewrite most of the model fitting, tuning, and evaluation code when switching from one model to another. 
The tidymodel approach unifies a wide range of models under a single pipeline where users can substitute different models while maintain the same pipeline.
This allows users to easily experiment with different machine learning models a lot easier.

**Constructing indices would also benefit from pipeline and embracing the aforementioned design philosophy.**

*more work here*

Both arguments are applicable to index construction.
Papers that proposing new indices always present a workflow diagram to illustrate how data is transformed into the final indices, but not much implementation has explicitly present this this step-by-step transformation in the code design.
Constructing indices also involves this working pattern of trials and errors on different options before finding the final solution. A tidy workflow will give researchers more flexibility to tune the steps as they wish and while maintain a readable and intuitive code base.

### 2022-10-27 paper/paper.tex
restructure section 2 data pipeline in R

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Demo arXiv template},
  pdfauthor={H. Sherry Zhang; Collaborators},
  pdfkeywords={spatio-temporal data, indices, data pipeline},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Demo arXiv template}
\author{
\textbf{H. Sherry Zhang}~\orcidlink{0000-0002-7122-1463}\\Department of
Econometrics and Business Statistics\\Monash University\\Melbourne,
VIC\\\href{mailto:huize.zhang@monash.edu}{huize.zhang@monash.edu}\\\\\\
\textbf{Collaborators}\\Department of Econometrics and Business
Statistics\\Monash University\\Melbourne, VIC\\}
\date{}
\begin{document}
\maketitle
\begin{abstract}
\begin{itemize}
\tightlist
\item
  Indices, useful, quantify severity, early monitoring,
\item
  A huge number of indices have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indices are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indices
  from spatio-temporal data,
\item
  This allows all the indices to be constructed through a uniform data
  pipeline and different indices to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indices into small building blocks shared by most of the indices.
\item
  The framework will be demonstrated with drought indices as examples,
  but appliable in general to environmental indices constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
spatio-temporal data \sep indices \sep 
data pipeline

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

The concept of pipeline (and pipe) is not a new concept to computing. It
refers to a set of data processing elements connected in series, where
the output of one element is the input of the next one. Wickham et al.
(2009) argues that whether made explicit or not, pipeline has to be
presented in every graphics program. The paper also argues that breaking
down graphic rendering into steps is beneficial for understand the
implementation and compare between different graphic systems.

\emph{The pipeline system would work partially because of an agreed data
structure (tidy data) and {[}same input and output format{]} - why the
pipeline approach would work.}

The tidyverse suite provides a collection of R package for conducting
data analysis. These tools directly operate on data, and users, even new
users, can easily read from the code on what operation is performed on
the data and what to expect as the output. All of these are built upon a
simple prior knowledge of the tidy data principle (Wickham 2014), which
prescribes three rules to structure tabular data: 1) Each variable forms
a column, 2) Each observation forms a row, and; 3) Each type of
observational unit forms a table.

Constructing indices also involves this working pattern of trials and
errors on different options before finding the final solution. A tidy
workflow will give researchers more flexibility to tune the steps as
they wish and while maintain a readable and intuitive code base.

To better understand the pipeline approach in statistic computing,
examples in interactive graphics and machine learning models are given.
In interactive graphics, the data pipeline is used to describe how the
data is processed to be rendered on the screen. Examples of early work
on building pipeline for interactive graphics includes Buja et al.
(1988), Sutherland et al. (2000), and Xie, Hofmann, and Cheng (2014).
Steps in these pipelines include non-linear transformation, variable
standardization, randomization and dimension reduction.

A more familiar example from tidymodels (Kuhn and Wickham 2020) presents
a set of packages for machine learning models following the tidyverse
principles (Wickham et al. 2019). The machine learning pipeline
constitutes data resampling, feature engineering, model fitting, model
tuning, and model evaluation each being responsible by a separate
package. Before tidymodel, analysts who wish to fit different models
need to learn the workflow of each individual algorithm and need to
rewrite most of the model fitting, tuning, and evaluation code when
switching from one model to another. The tidymodel approach unifies a
wide range of models under a single pipeline where users can substitute
different models while maintain the same pipeline. This allows users to
easily experiment with different machine learning models a lot easier.

Both arguments are applicable to index construction. Papers that
proposing new indices always present a workflow diagram to illustrate
how data is transformed into the final indices, but not much
implementation has explicitly present this this step-by-step
transformation in the code design.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

>>>>

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Demo arXiv template},
  pdfauthor={H. Sherry Zhang; Collaborators},
  pdfkeywords={indices, data pipeline, software design},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Demo arXiv template}
\author{
\textbf{H. Sherry Zhang}~\orcidlink{0000-0002-7122-1463}\\Department of
Econometrics and Business Statistics\\Monash University\\Melbourne,
VIC\\\href{mailto:huize.zhang@monash.edu}{huize.zhang@monash.edu}\\\\\\
\textbf{Collaborators}\\Department of Econometrics and Business
Statistics\\Monash University\\Melbourne, VIC\\}
\date{}
\begin{document}
\maketitle
\begin{abstract}
\begin{itemize}
\tightlist
\item
  Indices, useful, quantify severity, early monitoring,
\item
  A huge number of indices have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indices are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indices
  from spatio-temporal data,
\item
  This allows all the indices to be constructed through a uniform data
  pipeline and different indices to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indices into small building blocks shared by most of the indices.
\item
  The framework will be demonstrated with drought indices as examples,
  but appliable in general to environmental indices constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
indices \sep data pipeline \sep 
software design

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, interior hidden, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

@@

\textbf{Why you should care about pipeline}

The concept of pipeline (and pipe) is not a new concept to computing. It
refers to a set of data processing elements connected in series, where
the output of one element is the input of the next one. Wickham et al.
(2009) argues that whether made explicit or not, pipeline has to be
presented in every graphics program. The paper also argues that breaking
down graphic rendering into steps is beneficial for understand the
implementation and compare between different graphic systems. The
discussion on pipeline construction is well documented in early
interactive graphics software: Buja et al. (1988), Sutherland et al.
(2000), and Xie, Hofmann, and Cheng (2014) and their pipeline steps
include non-linear transformation, variable standardization,
randomization and dimension reduction.

\textbf{What is pipeline, its underlying software design philosophy, and
how these are reflected in R}

One of the most commonly known pipeline example is perhaps the Unix
pipeline where programs can be concatenated with \texttt{\textbar{}} to
flow the output from the last program into the next program. The
pipeline design avoid writing a complicated program to solve a complex
problem but building simple programs that do one thing well and work
well together. This design philosophy is also embraced in the R
community by the tidyverse principles. In tidyverse, the flow from the
previous program to the next is ensured through reusing the tidy data
format (Wickham 2014), which prescribes three rules on how to layout
tabular data. To solve a complicated data problem using tidyverse, one
typically builds the solution using a collection of functions and these
functions have a strong human centered design with intuitive naming and
easy to reference through autocomplete.

\textbf{A proper pipeline example in R: tidymodel}

A pipeline example in R that adopts the tidyverse principle is the
tidymodel suite (Kuhn and Wickham 2020) for building machine learning
models. The machine learning pipeline constitutes data resampling,
feature engineering, model fitting, model tuning, and model evaluation
each being responsible by a separate package. Before tidymodel, analysts
who wish to fit different models need to learn the workflow of each
individual algorithm and need to rewrite most of the model fitting,
tuning, and evaluation code when switching from one model to another.
The tidymodel approach unifies a wide range of models under a single
pipeline where users can substitute different models while maintain the
same pipeline. This allows users to easily experiment with different
machine learning models a lot easier.

\textbf{Constructing indices would also benefit from pipeline and
embracing the aforementioned design philosophy.}

\emph{more work here}

Both arguments are applicable to index construction. Papers that
proposing new indices always present a workflow diagram to illustrate
how data is transformed into the final indices, but not much
implementation has explicitly present this this step-by-step
transformation in the code design. Constructing indices also involves
this working pattern of trials and errors on different options before
finding the final solution. A tidy workflow will give researchers more
flexibility to tune the steps as they wish and while maintain a readable
and intuitive code base.

@@
